MSstatsPTM: Statistical relative quantification of post-translational modifications in bottom-up mass spectrometry-based proteomics

### 2022-07-13 main/MSstatsPTM_Main.tex
implementation converter section

The proposed methods are implemented in the open source R package $MSstatsPTM$, overviewed in~\figref{msstatsptm_design}. $MSstatsPTM$ is applicable to most experimental designs, label-free data acquisitions such as DDA, DIA, SRM and label-based data acquisitions such as TMT. $MSstatsPTM$ takes as input lists of identified and quantified spectral features, produced by spectral processing tools such as MaxQuant, Progenesis, or Spectronaut.  \todo{Could you rephrase the next sentence to be more specific. If someone is a tool developer, what exactly should they do?} To identify the local modification site we require the tool to identify where in the sequence the modification occurred.  The package includes converters for seamless integration with multiple such tools. \todo{Could you clarify what the converters do, especially for PTM? Do we discard peptides with multiple modification sites?}. For experimental designs that include separate runs enriched in modified peptides and global profiling runs,  conversion is performed separately for the runs enriched in modified peptides and for the global profiling runs. For experimental designs that include modified and unmodified peptides in a same run, \todo{what happens for the conversion?}

>>>>

The proposed methods are implemented in the open source R package $MSstatsPTM$, overviewed in~\figref{msstatsptm_design}. $MSstatsPTM$ is applicable to most experimental designs, label-free data acquisitions such as DDA, DIA, SRM and label-based data acquisitions such as TMT. $MSstatsPTM$ takes as input lists of identified and quantified spectral features, produced by spectral processing tools such as MaxQuant, Progenesis, or Spectronaut. We require the processing tool to identify the amino acid in the protein sequence where the modification occurred (i.e. the modification site). This will generally include the amino acid abbreviation, plus it's number in the protein sequence. For example, a modification on a serine, which was the 70th amino acid in the sequence would be marked ``S70''.  
Occasionally tools will only include the peptide sequence with the modified amino acid highlighted. If there is no way to determine where in the protein sequence the modification occurred, $MSstatsPTM$ includes functionality to do this manually, given the modified peptide sequence and a FASTA file with the entire protein sequence. The package includes converters for seamless integration with multiple such tools. The converters can optionally keep peptides with both single and multiple modification sites, or filter out only those with multiple sites. If multiple sites are kept, these are grouped separately from the singly modified peptide so as not to confound the effect. Additionally, peptides with oxidation can also be removed if the effect is believed to confound the modifications of interest. For experimental designs that include separate runs enriched in modified peptides and global profiling runs, conversion is performed separately for the runs enriched in modified peptides and for the global profiling runs.
